### 2023-12-24 paper.tex
Final cleanup

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Template for USENIX papers.
%
% History:
%
% - TEMPLATE for Usenix papers, specifically to meet requirements of
%   USENIX '05. originally a template for producing IEEE-format
%   articles using LaTeX. written by Matthew Ward, CS Department,
%   Worcester Polytechnic Institute. adapted by David Beazley for his
%   excellent SWIG paper in Proceedings, Tcl 96. turned into a
%   smartass generic template by De Clarke, with thanks to both the
%   above pioneers. Use at your own risk. Complaints to /dev/null.
%   Make it two column with no page numbering, default is 10 point.
%
% - Munged by Fred Douglis <douglis@research.att.com> 10/97 to
%   separate the .sty file from the LaTeX source template, so that
%   people can more easily include the .sty file into an existing
%   document. Also changed to more closely follow the style guidelines
%   as represented by the Word sample file.
%
% - Note that since 2010, USENIX does not require endnotes. If you
%   want foot of page notes, don't include the endnotes package in the
%   usepackage command, below.
% - This version uses the latex2e styles, not the very ancient 2.09
%   stuff.
%
% - Updated July 2018: Text block size changed from 6.5" to 7"
%
% - Updated Dec 2018 for ATC'19:
%
%   * Revised text to pass HotCRP's auto-formatting check, with
%     hotcrp.settings.submission_form.body_font_size=10pt, and
%     hotcrp.settings.submission_form.line_height=12pt
%
%   * Switched from \endnote-s to \footnote-s to match Usenix's policy.
%
%   * \section* => \begin{abstract} ... \end{abstract}
%
%   * Make template self-contained in terms of bibtex entires, to allow
%     this file to be compiled. (And changing refs style to 'plain'.)
%
%   * Make template self-contained in terms of figures, to
%     allow this file to be compiled.
%
%   * Added packages for hyperref, embedding fonts, and improving
%     appearance.
%
%   * Removed outdated text.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[letterpaper,twocolumn,10pt]{article}
\usepackage{usenix}

% to be able to draw some self-contained figs
\usepackage{tikz}
\usepackage{amsmath}
\usepackage{caption}

% inlined bib file
\usepackage{filecontents}

\usepackage{graphicx}
\graphicspath{ {./imgs/} }

%-------------------------------------------------------------------------------
\begin{filecontents}{\jobname.bib}
%-------------------------------------------------------------------------------
@InProceedings{hypsec,
  author =       {Li, Shih-Wei and Koh, John S. and Nieh, Jason},
  title =        {Protecting Cloud Virtual Machines from Commodity Hypervisor and Host Operating System Exploit},
  booktitle =    {Proceedings of the 28th USENIX Security Symposium},
  year =         2019,
  pages =        {1357--1374},
  note =         {\url{https://www.cs.columbia.edu/~nieh/pubs/security2019_hypsec.pdf}}
}
@InProceedings{sekvm,
  author =       {Li, Shih-Wei and Li, Xupeng and Gu, Ronghui and Nieh, Jason and Hui, John Z.},
  title =        {Formally Verified Memory Protection for a Commodity Multiprocessor Hypervisor},
  booktitle =    {Proceedings of the 30th USENIX Security Symposium.},
  year =         2021,
  pages =        {3953--3970},
  note =         {\url{https://www.usenix.org/system/files/sec21-li-shih-wei.pdf}}
}
@Misc{xuheng,
    author =        {Li, Xuheng},
    title =         {\text{Fall 23 Operating Systems 2 Homework}},
    year =          {2023},
    howpublished =  {\url{https://xuhengli.notion.site}},
    note =          {Accessed: 2023-12-21}
}
@Misc{arm,
    author =        {ARM Developer Documentation},
    title =         {Learn the architecture - AArch64 virtualization},
    howpublished =  {\url{https://developer.arm.com/documentation/102142/0100/Stage-2-translation}},
    note =          {Accessed: 2023-12-21}
}
@Misc{vmlinux.lds.S,
    author =        {XuhengLi (GitHub username)},
    title =         {vmlinux.lds.S in osdi23-paper114-sekvm},
    howpublished =  {\url{https://github.com/columbia/osdi23-paper114-sekvm/blob/ae/arch/arm64/kernel/vmlinux.lds.S}},
    note =          {Accessed: 2023-12-21}
}
@Misc{kernel-pgtable.h,
    author =        {XuhengLi (GitHub username)},
    title =         {kernel-pagetable.h in osdi23-paper114-sekvm},
    howpublished =  {\url{https://github.com/columbia/osdi23-paper114-sekvm/blob/ae/arch/arm64/include/asm/kernel-pgtable.h}},
    note =          {Accessed: 2023-12-21}
}
@Misc{el1.c,
    author =        {XuhengLi (GitHub username)},
    title =         {el1.c in osdi23-paper114-sekvm},
    howpublished =  {\url{https://github.com/columbia/osdi23-paper114-sekvm/blob/ae/arch/arm64/hypsec_proved/el1.c}},
    note =          {Accessed: 2023-12-21}
}
@Misc{kerneldocs,
    author =        {kernel.org},
    title =         {Chapter 6: Physical Page Allocation},
    howpublished =  {\url{https://www.kernel.org/doc/gorman/html/understand/understand009.html}},
    note =          {Accessed: 2023-12-22}
}
@Misc{page_alloc.c,
    author =        {Elixir Bootlin},
    title =         {page\_alloc.c in Linux v5.4.55},
    howpublished =  {\url{https://elixir.bootlin.com/linux/v5.4.55/source/mm/page_alloc.c#L4718}},
    note =          {Accessed: 2023-12-22}
}
@Misc{mmzone.h,
    author =        {Elixir Bootlin},
    title =         {mmzone.h in Linux v5.4.55},
    howpublished =  {\url{https://elixir.bootlin.com/linux/v5.4.55/source/include/linux/mmzone.h#L27}},
    note =          {Accessed: 2023-12-22}
}
@Misc{el2.c,
    author =        {XuhengLi (GitHub username)},
    title =         {el2.c in osdi23-paper114-sekvm},
    howpublished =  {\url{https://github.com/columbia/osdi23-paper114-sekvm/blob/ae/arch/arm64/hypsec_proved/el2.c}},
    note =          {Accessed: 2023-12-22}
}
@Misc{hypsec_host.h,
    author =        {XuhengLi (GitHub username)},
    title =         {hypsec\_host.h in osdi23-paper114-sekvm},
    howpublished =  {\url{https://github.com/columbia/osdi23-paper114-sekvm/blob/ae/arch/arm64/include/asm/hypsec_host.h}},
    note =          {Accessed: 2023-12-22}
}
\end{filecontents}

% %-------------------------------------------------------------------------------
\begin{document}
% %-------------------------------------------------------------------------------

%don't want date printed
\date{}

% make title bold and 14 pt font (Latex default is non-bold, 16 pt)
\title{\Large \bf FlexAlloc: Dynamic Memory Partitioning\\
    between the Corevisor and Hostvisor in SeKVM}

%for single author (just remove % characters)
\author{
{\rm Matthew Nelson}\\
Columbia University
\and
{\rm Ryan Wee}\\
Columbia University
\and
{\rm Zeheng Yang}\\
Columbia University
} % end author

\maketitle

%-------------------------------------------------------------------------------
\begin{abstract}
%-------------------------------------------------------------------------------
SeKVM is a hypervisor for ARM processors that protects the memory of guest
virtual machines (VMs) from an untrusted host machine. It comprises a
corevisor that runs at EL2, and a hostvisor that runs at EL1
to leverage the functionality of the untrusted host kernel. However, SeKVM
statically partitions memory between the corevisor and hostvisor at compile
time. This leads to memory wastage when the number of VMs is low, and an
inability to support the desired number of VMs when the number of VMs is high.
In this paper, we present FlexAlloc, an extension of SeKVM that allows dynamic
runtime partitioning of memory between the corevisor and the hostvisor. In
particular, the memory needed for the stage-2 page tables of a guest VM is
only allocated to the corevisor when a VM is created. This memory is then
returned to the hostvisor when a VM is destroyed. Existing implementions
of \textit{alloc\_pages} do not support allocating a contiguous region of memory
equal in size to the memory needed for a VM's stage-2 page tables. Hence,
FlexAlloc also modifies the iterators used to accesss a VM's stage-2 page
tables, so that the memory used for a VM's stage-2 page tables can be composed
of noncontiguous regions. Our work ensures that a machine running SeKVM uses
memory more efficiently, and also allows such machines to support a larger
number of guest VMs. This allows public cloud providers and organizations
running their own VMs to scale up their operations at a lower cost.
\end{abstract}

%-------------------------------------------------------------------------------
\section{Introduction and Background}
%-------------------------------------------------------------------------------

Virtual machines (VMs) have become increasingly popular over the past few
decades. In particular, they improve utilization of existing hardware by
allowing users to run multiple unmodified operating systems on a single machine.
A key concern when it comes to VMs is isolation. A single machine may be used to
run multiple VMs created by different users who generally want their data
to be protected from one another. As a result, hypervisors generally maintain
some sort of secure boundary between guest VMs. However, another important
aspect of isolation is protecting guest VMs from the host machine. A hypervisor
that fully trusts its host machine is vulnerable to malicious actors who
gain control of the host operating system itself. This is dangerous because
the host machine could be used to steal confidential corporate data or sensitive
personal information.

SeKVM is a formally verified hypervisor for ARM that isolates KVM's trusted computing
base into a small core~\cite{sekvm}. It builds on the previous work of HypSec,
which is a hypervisor that protects VMs from the host machine~\cite{hypsec}.
HypSec consists of two layers. The first layer is an untrusted
\textit{hostvisor}, which operates at EL1 together with the host operating
system. The hostvisor leverages many of the data structures and functions
built into the host operating system for VM management. The second layer is a
trusted \textit{corevisor}, which operates at the more privileged EL2 level.
This corevisor is responsible for isolating VMs from each other and from the
host machine. SeKVM builds on HypSec's design, with the main difference being
that its trusted computing base is formally verified. In SeKVM, the two layers
are termed \textit{KServ} and \textit{KCore} respectively. However, in this
paper, we use the terms hostvisor and corevisor to remain
consistent with the underlying HypSec kernel source code.

SeKVM protects guest VMs in three main areas: CPU, memory, and I/O. In the
area of memory, SeKVM leverages ARM hardware support for virtualization. ARM
allows hypervisors to control memory access using stage-2
translation~\cite{arm}. In particular, the guest operating systems and host
operating system each think they are using page tables to map a virtual address space to the
physical address space. However, in reality, they only control the mapping
of virtual addresses to intermediate physical addresses. When translating
virtual addresses, the MMU starts by using these page tables to translate
virtual addresses into intermediate physical addresses. However, it then
uses a second set of page tables to translate
intermediate phyiscal addresses into the actual machine physical addresses.
This second set of page tables, called stage-2 page tables, are controlled by
the hypervisor running at EL2. In the context of SeKVM, the stage-2 page tables
are managed by the corevisor. When the host operating system tries to
access some physical memory, the corevisor first checks that this
physical memory does not belong to a guest VM.

The hostvisor cannot be allowed to access the memory used to store
the stage-2 page tables of guest VMs. Otherwise, it could simply modify these
stage-2 page tables so that guest VMs write their data to memory that can be
accessed and manipulated by the hostvisor. In other words, the memory used for
these stage-2 page tables must `belong' to the corevisor. In SeKVM,
the corevisor is statically allocated this memory on bootup. This is done by
a linker script, which allocates some
contiguous region of memory for the stage-2 page tables and demarcates it using the labels \textit{stage2\_pgs\_start}
and \textit{stage2\_pgs\_end}~\cite{vmlinux.lds.S}. The size of this memory region is
determined by the macro \textit{STAGE2\_PAGES\_SIZE}~\cite{kernel-pgtable.h}. In
particular, this memory region is large enough to store the stage-2 page table
for the corevisor, the stage-2 page table for the hostvisor,
and the stage-2 page tables for sixteen guest VMs. The corevisor and hostvisor are
each allocated sixteen 2M pages for their stage-2 page tables, while each guest VM
is allocated four 2M pages. Altogether, the corevisor is given 96 2M
pages for all of these stage-2 page tables. During bootup, the hostvisor
allocates the beginning of this memory region to the corevisor and itself. It then
initializes the corevisor's VM-management data structure such that the remainder of this stage-2 page table memory region
is evenly divided between each of the sixteen possible guest VMs. In particular, the hostvisor sets
\textit{el2\_data->vm\_info[i].page\_pool\_start =
pool\_start + (STAGE2\_VM\_POOL\_SIZE * (i - 1))} for each possible VMID~\cite{el1.c}.
Here, \textit{vm\_info[i].page\_pool\_start} marks the start of the stage-2
page table for the \textit{i}th VM.

The key observation here is that this partitioning of memory is done statically
at compile time. In other words, the hostvisor cannot reclaim the memory it has
allocated to the corevisor. Similarly, the corevisor cannot ask for more memory
from the hostvisor. This is in contrast to the dynamic partitioning of memory
between the hostvisor and guest VMs. Guest VMs can request memory from the
hostvisor via a page fault, and the hostvisor can reclaim memory from guest VMs
using ballooning~\cite{hypsec}. The static partitioning of memory between the
hostvisor and the corevisor is undesirable for two reasons:

\begin{itemize}
    \item First, unused memory cannot be reclaimed from the corevisor. Say
    the number of VMs being run is $n$, such that $n < 16$. Then there will be
    $(16 - n) * 4$ 2M pages reserved for the stage-2 page tables of guest VMs that
    do not exist. Dynamic partitioning of memory would allow this memory to be reclaimed by the hostvisor, to
    alleviate memory pressure in the host. This memory could also be given to guest VMs, to
    alleviate memory pressure in these VMs. Giving VMs more memory would improve
    VM performance, because less data would need to be moved to the swap partition.
    \item Second, there is a hard upper limit on the number of VMs that can
    be supported by SeKVM. Users cannot add VMs above the limit of sixteen,
    because there will be no memory in the corevisor to support the stage-2
    page tables of these VMs. Users could always choose to statically allocate
    even more memory to the corevisor upon bootup, but this would exacerbate
    the problem of unused memory mentioned above.
\end{itemize}

Static allocation of memory to the corevisor is not limited to memory for the
stage-2 page tables of each VM. In addition to this, SeKVM also statically
allocates memory to the corevisor to store the metadata of each VM. The same linker script
mentioned above reserves another contiguous region of memory equal in size to 32
2M pages, and demarcates it using \textit{el2\_data\_start}
and \textit{el2\_data\_end}~\cite{vmlinux.lds.S}. This memory is given to the
corevisor on bootup. Part of this memory is used to
store an array of \textit{struct el2\_vm\_info} objects, where each object is
used to store the metadata of each guest VM. SeKVM could potentially also benefit
from dynamic allocation of the memory used to store this VM metadata, where
this memory is only given to the corevisor upon VM creation and returned to the
hostvisor upon VM deletion. However,
we choose not to focus on this because the size of each \textit{struct el2\_vm\_info} object
is 768 bytes. The total size of all of the \textit{struct el2\_vm\_info} objects
is 12288 bytes, which is less than four 4K pages in total. As a result, the
problem of unused memory is not as significant, since we waste at most four
4K pages of memory. In any case, memory can only be allocated at the
page granularity. Hence, in the case where we want to create a new VM and the new \textit{struct el2\_vm\_info} object
cannot be placed into the pages that have already been dynamically allocated to the corevisor, we will
inevitably end up having to allocate an entirely new 4K page. The majority of this new 4K page would then remain unused, until another VM is created. This
is not the case with VM stage-2 page tables, since the size of the stage-2 page
table for each VM is nicely page-aligned. With the \textit{struct el2\_vm\_info} objects, the problem of a hard upper limit
is also not as significant, since we could easily make the array larger without
incurring much memory overhead.

Although the \textit{struct el2\_vm\_info} objects have a much smaller memory footprint than the stage-2 page tables,
we designed and partially implemented a fully-generalized dynamic allocation interface that
could be used to support dynamic allocation for all metadata structs. We call this the
Dynamic Page Manager (DPM). DPM handles all of the work required to receive memory from and return
memory to the hostvisor, allowing other ``services'' in EL2 to exchange memory with DPM using
a very similiar interface to the \textit{alloc\_pages} and \textit{free\_pages} interface available in EL1. We describe
this interface in detail in Section 4. However, the next section focuses
exclusively on the changes made to support dynamic allocation of the memory
needed for VM stage-2 page tables.

%-------------------------------------------------------------------------------
\section{Design and Implementation}
%-------------------------------------------------------------------------------

All of the source code described below can be found at \url{https://github.com/nelsonm2991/host-sekvm-project}.
Our implementation of FlexAlloc is based on SeKVM, which is in turn based on Linux v5.4.55.

\subsection{Modifying the page table iterators}

As described above, the original implementation of SeKVM statically
allocates sixteen contiguous 8M regions to the corevisor (Figure \ref{fig:sekvmvmregions}). When a VM is created,
the corevisor uses one of these contiguous 8M regions for the VM's stage-2
page table. To implement a system of dynamic allocation, we want the hostvisor
to dynamically reserve an 8M region at runtime when a VM is created, and then
pass this memory region to the corevisor. The kernel macro \textit{alloc\_pages}
is responsible for allocating a large number of contiguous physical pages~\cite{kerneldocs}.
Examining the implementation of this macro reveals that it eventually
calls \textit{\_\_alloc\_pages\_nodemask}, which checks that the order requested
is strictly less than \textit{MAX\_ORDER}~\cite{page_alloc.c}. \textit{MAX\_ORDER}
is in turn defined to be 11~\cite{mmzone.h}. This means that we cannot allocate
a contiguous region of memory larger than $2^{10} * 4K = 4M$. This
limit is set by the kernel because \textit{alloc\_pages} uses the buddy allocator,
which recursively splits available memory into half to find the best possible position for
a contiguous region of memory~\cite{kerneldocs}. However, this also makes it
undesirable to continuously allocate huge contiguous regions of memory.

\begin{figure}[h!tbp]
\centering
\captionsetup{justification=centering}
\includegraphics{sekvm_vm_region}
\caption{SeKVM: Layout of Statically-Allocated\\Region for VM Stage-2 Page Tables}
\label{fig:sekvmvmregions}
\end{figure}

\bigskip To get around this limitation, there are two possible approaches. The first approach
is to halve the size of the VM's stage-2 page table, so that it can fit within
a single contiguous 4M region of memory. Alternatively, we can fragment the VM's stage-2
page table such that it can be composed of multiple non-contiguous regions of memory, whose sizes in total add up to 8M. We
choose to implement the latter approach to avoid degrading VM performance, and to avoid
breaking the current implementation of the buddy allocator. In
particular, we fragment the VM's stage-2 page table so that it can be composed
of eight distinct 1M regions of memory. Each of these 1M regions must be contiguous,
but each region is distinct and need not be allocated contiguously with the
other 1M regions. As a result, the hostvisor can easily use \textit{alloc\_pages}
to allocate the memory needed for a guest VM's stage-2 page tables.

We implement this fragmentation by manipulating the iterators that provide the next useable
page for entries at a given level of the page table. In the original SeKVM implementation, the pages containing entries from some level of the stage-2 page table are
kept contiguous to one another, and are never intermixed with pages containing entries from other levels. Hence, there are distinct iterators for each level.
Each iterator starts at a particular offset into the VM's statically-allocated 8M stage-2
page table region, and increments each time we need a new page for a new entry at that level. In particular, the PGD iterator operates over a \textit{16 * PAGE\_SIZE} pool, the PUD
iterator operates over a \textit{SZ\_2M - (16 * PAGE\_SIZE)} pool, and the PMD iterator operates over a \textit{3 * SZ\_2M} pool.
Each of these pools must be contiguous, and the three pools must be contiguous relative to one another.

The naming semantics of the iterators at each level can be confusing, so we feel they are worth
clarifying here. The iterator for level $n$ allocates pages for entries in level $n$ to point to.
Hence, the iterator for level $n$ points to the pool used to allocate
entries belonging to level $n-1$. For instance, the PGD iterator allocates pages
for PGD entries to point to. In other words, the PGD iterator actually allocates
pages used to hold PUD entries. Hence, the PGD iterator is initialized to point to the start of the PUD pool.

Our design eliminates the need for all of these pools to be contiguous to
one another, and enables fragmentation within each pool itself. To achieve this goal,
we experiment with two distinct implementations. As mentioned above, the SeKVM implementation
structures iterators such that a page with PUD entries will never be sandwiched between a page with PMD entries.
In our first implementation, we ignore this semantic. In particular, we create a single,
generalized iterator shared across all levels of the stage-2 page table. This
iterator traverses all eight 1M regions. When it exhausts the space in a 1M region, it moves
to the start of the next 1M region. However, we eventually decided to stop development of this
version before it was fully functional, in favor of version two which more closely resembles recent changes made to SeKVM's
stage-2 page table allocation scheme.

In our second implementation, we enforce the semantic that pages at each specific level are
allocated from seperate pools, and thus are never intermixed with one another. To accomplish
this while still maintaining the same pool of non-contiguous 1MB regions,
we split the single, generalized iterator into nine distinct iterators. 
One PGD iterator is used to allocate pages from the PUD pool, two PUD iterators are used to allocate pages from the PMD pool,
and six PMD iterators are used to allocate pages from the PTE pool. There are two key things to
note here. First, we retain the existing scheme partitioning the stage-2 page table into
three distinct pools: a PUD pool, a PMD pool, and a PTE pool.
For instance, the PUD pool is still sized at \textit{16 * PAGE\_SIZE}. The
second key thing to note is that our iterators never cross the boundary
between two distinct 1M regions. Instead, if a pool needs to be split across
two or more 1M regions, there is one distinct iterator for each 1M region
covered by the pool. As a result, each of these iterators covers a contiguous subregion of the pool.
Allocations from that pool begin by using the first iterator. When the first
subregion of that pool is exhausted, allocations move on to use the next iterator.
Essentially, when a page is allocated from a given pool, the correct iterator is chosen based on
the pool in question and whether or not the preceding iterators for that pool have been exhausted.

% We do that by modfifying , the FlexAlloc
% iterators operate over a pool of eight non-contiguous 1MB regions. The FlexAlloc iterators
% first exhaust all entries from a given 1MB region. Once the current 1MB region is exhausted, FlexAlloc
% will jump the iterators to the next 1MB region in the pool. Subsequent traversals of the
% stage-2 page tables with the alloc flag enabled will pull pages from the seperate, second
% non-contiguous 1M region. By removing the requirement for contiguous 8MB regions, FlexAlloc
% enables reliable dynamic allocation of stage-2 page table regions.

\bigskip
\begin{figure}[h!tbp]
\centering
\captionsetup{justification=centering}
\includegraphics{firstmbregion}
\caption{FlexAlloc: Metadata Setup\\for First 1MB Region}
\label{fig:firstmbregion}
\end{figure}

\begin{figure}[h!tbp]
\centering
\captionsetup{justification=centering}
\includegraphics{otherregions}
\caption{FlexAlloc: Metadata Setup\\for Second to Eighth 1MB Regions}
\label{fig:otherregions}
\end{figure}

In particular, the PUD pool is in the middle of the first 1M region, and there is a single PGD iterator
pointing to the start of that pool (Figure \ref{fig:firstmbregion}). The PMD pool consumes the end of the first 1M region
and the entire second 1M region. Hence, the first PUD iterator points to an address in
the first 1M region, and the second PUD Iterator points to the start of the
second 1M region (Figures \ref{fig:firstmbregion}, \ref{fig:otherregions}). Finally, the PTE pool consumes the third to eighth 1M regions.
As a result, there is a distinct PMD iterator for each of these six regions
(Figure \ref{fig:otherregions}). The code below illustrates how allocations from a given pool are made using
the first possible iterator whose pool subregion has not
been exhausted:

\begin{verbatim}
static u64 inline get_pmd_next(u32 vmid) {
    ...
    // pool_index: available iterator
    // we are using
    pool_index =
      el2_data->vm_info[vmid].pte_pool_index;
    // pool_start: beginning of the region
    // the pool_index corresponds to
    pool_start =
      el2_data->vm_info[vmid].
      pte_pool_starts[pool_index];
    // used_pages: number of pages in this
    // region we have exhausted
    used_pages = el2_data->vm_info[vmid]
      .pte_used_pages_vm[pool_index];

    // If we've used up all the pages and we
    // have another region, then use the new
    // region.
    if (used_pages == PTE_ITER_MAX_PAGES &&
        pool_index + 1 <
        PTE_USED_ITER_COUNT) {
        pool_index += 1;
        el2_data->vm_info[vmid].
          pte_pool_index = pool_index;

        // Set the pool_start to the next
        // available PTE iterator.
        pool_start = el2_data->vm_info[vmid].
          pte_pool_starts[pool_index];
        used_pages = el2_data->vm_info[vmid].
          pte_used_pages_vm[pool_index];
    }

    // Make sure we have not exhausted all
    // pages across all regions.
    if (used_pages == PTE_ITER_MAX_PAGES &&
            pool_index ==
            PTE_USED_ITER_COUNT - 1) {
        __hyp_panic();
    }

    // Return the physical address of the
    // next page that this iterator can
    // allocate.
    return pool_start +
      (used_pages * PAGE_SIZE);
}
\end{verbatim}

\subsection{Allocating memory to the corevisor}

FlexAlloc dynamically allocates eight 1M memory regions to the corevisor when a
VM is created. It does so by passing the base addresses of these eight 1M
regions as arguments to the \textit{HVC\_REGISTER\_KVM} hypercall. An alternative
implementation that we considered was creating a new \textit{HVC\_PASS\_MEM} hypercall to pass these
1M regions to the corevisor, and getting the hostvisor to call \textit{HVC\_PASS\_MEM}
before calling \textit{HVC\_REGISTER\_KVM}. However, we ultimately choose to
discard this alternative design because it leads to a more complicated
implementation. In particular, the corevisor would have to verify a one-to-one
correspondence between calls to \textit{HVC\_PASS\_MEM} and \textit{HVC\_REGISTER\_KVM}.
It would also have to handle malicious hostvisors that try to call
\textit{HVC\_REGISTER\_KVM} without calling \textit{HVC\_PASS\_MEM}. Morevoer,
adding an extra hypercall increases latency because it leads to more frequent
world-switches between the hostvisor running in EL1 and the corevisor running in EL2.

The \textit{HVC\_REGISTER\_KVM} hypercall eventually calls \textit{register\_kvm}
~\cite{el2.c}. Note that at this point we are in the corevisor, running at EL2.
Within this function, we preprocess the eight 1M regions to ensure that the
calls to \textit{alloc\_pages} in the hostvisor were successful, and to verify
that the memory regions provided actually belong to the hostvisor. We then set
the ownership of these memory regions to the corevisor, so that the host can
no longer access or manipulate the guest VM's stage-2 page tables. In particular,
we do the following:

\begin{verbatim}
for (i = 0; i < 8; ++i) {
    addr = page_starts[i];
    end = page_starts[i] + SZ_1M;

    /* ... */

    if (addr == 0) {
        // Hostvisor's alloc_pages() failed.
        __hyp_panic();
    }

    // Change the ownership of these pages
    // to the corevisor.
    do {
        index = get_s2_page_index(addr);
        owner = get_s2_page_vmid(index);

        if (owner != HOSTVISOR) {
            // The hostvisor shouldn't be
            // giving memory that belongs
            // to the corevisor or to a VM.
            __hyp_panic();
        }

        set_s2_page_vmid(index, COREVISOR);
        addr += PAGE_SIZE;
        ++page_cnt;
    } while (addr < end);
}
\end{verbatim}

\noindent We then build the VM's stage-2 page table using these eight 1M regions, by
populating the relevant fields in the \textit{struct el2\_data} object used
by the corevisor for VM management.

\subsection{Reclaiming memory from the corevisor}

When a VM is destroyed, FlexAlloc reclaims the eight 1M memory regions that were
previously allocated to the corevisor for the VM's stage-2 page table. Returning
this memory to the hostvisor ensures that it does not remain unused in corevisor
memory. Unlike the case of VM creation, there is no hypercall made directly to
the corevisor when a VM is destroyed. Hence, we create a new \textit{HVC\_DESTROY\_KVM}
hypercall that mirrors the \textit{HVC\_REGISTER\_KVM} hypercall used during VM
creation. In particular, on ARM architectures, if the \textit{CONFIG\_VERIFIED\_KVM} macro is defined,
we call \textit{hypsec\_arch\_destroy\_vm}. This function in turn calls \textit{hypsec\_destroy\_kvm},
which ultimately invokes the \textit{HVC\_REGISTER\_KVM} hypercall with the corresponding
VMID. We do not pass the base addresses of the eight 1M memory regions as arguments
to this hypercall, since the corevisor can access them directly by reading the
relevant fields of the \textit{struct el2\_data} object. Once we are in the
corevisor running in EL2, we return the eight memory regions to the hostvisor.
We then return to the hostvisor running in EL1, which can then call \textit{\_\_free\_pages}
on each of the eight 1M regions.

%-------------------------------------------------------------------------------
\section{Evaluation}
%-------------------------------------------------------------------------------

We evaluate FlexAlloc by running it on a QEMU-emulated Raspberry Pi 4. In
particular, we run VMWare Fusion Professional Version 13.0.0 on a 2018 x86-64 Macbook
Pro running MacOS Ventura 13.6.1. Using VMWare, we create a VM
with 4 processor cores and 12 GB memory running Ubuntu 22.04.3 LTS. On this
VM, we run QEMU 6.2.0 to emulate a Raspberry Pi 4. This emulated Raspberry
Pi 4 serves as our host machine, on which we run an SeKVM with the FlexAlloc
extension. Our guest VMs run vanilla Linux v5.4.

Using FlexAlloc, we boot up a guest VM and verify that it is able to start up
without any kernel panics. We use \textit{print\_string} to verify that we
are passing eight distinct 1M regions to the corevisor when the VM is created.
We test a case where the 1M regions are not contiguous, and verify that our
modified page table iterators are able to successfully handle a VM stage-2 page
table that is non-contiguous. We also shut down the guest VM and verify that
all eight 1M regions are returned to the hostvisor. We compare the base address
of each region allocated during VM creation against the base address of each
region reclaimed during VM shutdown, and verify that they match each other.

In order to verify that the hostvisor will generally be able to find eight distinct contiguous 1M regions when booting a VM, we evaluate how many times we can successfully call
\textit{alloc\_pages(1M)} before a null pointer is returned. To reproduce our tests,
simply compile \textit{host-sekvm-project} with the \textit{ALLOC\_STRESS\_TEST} macro defined.
Across our three trials, after booting host SeKVM with the FlexAlloc extension,
we were able to achieve between 7049 and 7054 successful 1M allocations
from calls to \textit{alloc\_pages(1M)}. Since each VM only needs eight successful allocations for its stage-2 page table,
we believe that the upper bound on the number of VMs we can create before running into \textit{alloc\_pages(1M)}
failures is sufficiently high.

We also verify the changes in the corevisor's memory footprint by checking the number of
pages owned by the corevisor, using the \textit{s2\_pages} array in \textit{el2\_data} at key points
when a VM is created and destroyed. On a host running SeKVM without the
FlexAlloc extension, we observe that the corevisor owns 65549 pages before booting a VM,
while running a VM, and after destroying a VM. On a host running SeKVM with the FlexAlloc
extension, we observe that the Corevisor owns 32768 pages before VM boot, 34817 pages
during the life of the VM, and 32768 pages after destroying the VM.
Note that FlexAlloc also drastically reduces the size of the statically-allocated \textit{STAGE2\_PAGES\_SIZE} region in the linker script,
since it now only needs to contain the stage-2 page tables of the hostvisor and the corevisor.
These findings show that the corevisor in FlexAlloc is able to successfully receive, use, and return pages
given to it by the hostvisor for constructing a VM's stage-2 page table.

%-------------------------------------------------------------------------------
\section{Shortcomings and Future Work}
%-------------------------------------------------------------------------------

\subsection{Shortcomings in design}

One key shortcoming of FlexAlloc is that it is not formally verified. In
this sense, FlexAlloc is less like SeKVM and more like HypSec. It provides isolation guarantees
in principle, but cannot be logically proven to be secure. In particular, there
are a few major flaws in FlexAlloc when it comes to isolating guest VMs from
the host machine. For instance, the current implementation of FlexAlloc
does not use \textit{memset} to zero out the memory given to it by the hostvisor
for the VM's stage-2 page table. This presents a danger if the MMU tries to
read from a page table entry that should not actually exist, but that has
previously been written into the memory region by the hostvisor. We also occasionally
observed stalls when shutting down VMs and booting new ones, which we
believe is caused by our failure to \textit{memset} the pages to zero. For example,
if we shut down a VM and boot up a new one, the second VM's stage-2 page table could contain pages previously
used by the first VM's stage-2 page table. These pages could contain non-zero data,
leading to undefined behavior when interpreted as page table entries.

Another flaw is that
FlexAlloc does not verify that the eight 1M regions are actually distinct and
do not overlap with one another. Hence, FlexAlloc still fundamentally relies on
a non-malicious hostvisor, and would require further patches to be fully secure.
To fix this, FlexAlloc could simply have an additional check verifying that all of the
physical addresses received from the host are unique, 1M-aligned, and non-overlapping.

Finally, our design does not fully implement
dynamic memory partitioning between the corevisor and hostvisor, outside of VM stage-2 page tables. As mentioned in
the introduction, another area in which corevisor memory scales with the number
of VMs is in maintaining \textit{struct el2\_vm\_info} objects that record the
metadata associated with each guest VM. Currently, we still statically allocate
a fixed-size array of \textit{struct el2\_vm\_info} objects in corevisor
memory ~\cite{hypsec_host.h}. This essentially preserves the hard cap on the
number of guest VMs that can be run. An easy fix would be to keep this array
as a statically-allocated array, while increasing the number of elements. Since each
\textit{struct el2\_vm\_info} object is only 768 bytes, this would not waste a
significant amount of memory even if fewer than sixteen VMs are created. Moreover,
each VM also requires a \textit{struct kvm}. Hence, to lift the sixteen-VM cap present
in SeKVM, we would need to either dynamically allocate the memory for this struct,
or increase the size of the statically-allocated region for VM metadata.
The latter approach is probably reasonable.
Since the changes to the VM stage-2 page tables already decrease the size
of the statically-allocated region by 128 MB, re-inflating this region to accomodate 32+ VMs would still result in substantial memory savings. 
The former approach is possible as well. In the subsection on Future Work, we propose DPM as a potential solution to
the issue of bringing dynamic allocation to EL2 metadata structures that are not page-aligned.

\subsection{Shortcomings in evaluation}

We also do not fully evaluate our implementation. For one, we do not stress-test our implementation by running a large number
of VMs. In addition, we do not verify whether fragmenting the stage-2 page table results
in any degradation in VM performance. Given the limitations of our test environment, we
only considered metrics related to the corevisor's memory footprint, and did not investigate performance metrics related to execution time. Ideally, we would run a memory-intensive
benchmark on our guest VMs to verify that our iterators work properly under a
high memory load across multiple VMs.

\subsection{Future work}

A possible area of future work is exploring other allocators designed for
allocating large chunks of contiguous memory. We fragment the VM stage-2 page tables
into eight distinct 1M pages, because \textit{alloc\_pages} does not allow us
to allocate contiguous regions larger than 4M. We also wanted to be reasonably certain
that the size of our regions would not impeed the host's ability to launch VMs. It might be possile to avoid this
by looking at other Linux memory allocators explicitly designed for the purpose
of large contiguous allocations and adding such an allocator to the host.

\subsection{Dynamic Page Manager}

\subsubsection{Overview}
To enable dynamic allocation of EL2 memory in general beyond just the scope of
a VM's stage-2 page tables, we also designed but did not fully implement
a dynamic page manager (DPM). DPM is a system that exposes an interface for other ``services'' within EL2 to
utilize dynamic allocation for their memory regions, while hiding the implementation of actually receiving memory from
and returning memory to the host. This would be another interesting area of future work. Please
see the \textit{dpm} branch of the linked repository for a partial DPM implementation.

By implementing an interface for EL2 to dynamically ``request'' and ``return'' arbitrary
regions of contiguous memory from EL1, the corevisor would no longer be limited to using statically-allocated, fixed-size regions.
Instead, other services within EL2 could use DPM to dynamically request regions of contiguous memory.
These services would be responsible for managing the contents of these pages once DPM
allocates these pages to that specific service. The service would be required to signal DPM
when that contiguous region is no longer in use, using an interface that is similar to the kmalloc and kfree model.

DPM operates at the contiguous region granularity rather than the page granularity because
a lot of the existing SeKVM code assumes contiguous regions. If DPM were to instead piece
together pages to form fragmented regions, all of this other code would break. Thus DPM
can actually be thought of as a memory-region manager rather than a page manager. DPM is structured
in such a way that a second page-granularity version of DPM could operate alongside the
region-granularity version of DPM, thus allowing some services to remain dependent on
contiguous regions while others adapt to use separated non-contiguous regions, in order to achieve better memory
utilization by EL2.

\subsubsection{Interface}
Our design of DPM exposes the following additional hypercalls to the
hostvisor running at EL1:

\begin{itemize}
\item Allocating pages to EL2:\\
\textit{hvc\_give\_pages(base\_phys\_addr\_of\_region, size\_of\_region)}.
Note that the size must be page-aligned.
\item Reclaiming pages from EL2:\\
\textit{hvc\_reclaim\_pages(base\_phys\_addr\_of\_output\_struct)}.
The output struct is provided by the hostvisor, and should contain an array of
physical addresses. Services at EL2 should record the physical addresses
of  regions being handed back to the hostvisor in the output
struct. The hostvisor can then read from the output struct and free the relevant pages.
Every physical address that EL2 records in this output struct is considered
freed and no longer managed by DPM.
\end{itemize}

\noindent Our design of DPM also exposes the following interface to services running
at EL2 that want dynamically-allocated memory:
\begin{itemize}
\item Requesting a contiguous region of a given size from the DPM pool:
\textit{dpm\_alloc\_region(size)}.
	Returns an address to the region if one was available in the pool,
    and 0 to indicate an error which the requesting service must propagate back to EL1.
\item
\textit{dpm\_free\_region(starting\_region\_addr, optional: size)}:
Signals that a previously-allocated region can be returned to the DPM pool.
\end{itemize}

\subsubsection{Metadata}

DPM manages regions of pages using a circular linked list (Figure \ref{fig:dpmmetadata}). When an EL2 service requests
a region of a specific size from DPM, DPM will search this list for an available match.
If no matches are found, DPM returns an error indicating to the EL2 service that it must
propogate the error back to the host. Upon receiving the error, the host will make an
additional hypercall that passes a new region of some size to DPM. After restarting the original
hypercall, DPM will successfully find a region to hand to the EL2 service. The code
below illustrates the data structures used to store this metadata.

\begin{verbatim}
// DPM Overall Metadata
// Contained in struct el2_data
struct dpm {
    // Address to the first
    // dpm_region page
    u64 base_page_addr;

    // Number of regions currently in pool
    u64 region_count;

    // Number of pages currently in pool
    u64 page_count;     
};

// DPM Region-Specific Metadata
struct dpm_region {
    // Addr of the beginning of the first
    // page in the region
    u64 start_addr;

    // Number of pages in region
    u64 page_count;

    // DXXX XXXA, where A is allocated flag,
    // D is DPM flag meaning that region
    // holds DPM metadata, X is available bit
    u8 flags;

    struct dpm_region* prev;
    struct dpm_region* next;
};
\end{verbatim}

\begin{figure}[h!tbp]
\centering
\captionsetup{justification=centering}
\includegraphics{dpmmetadata}
\caption{FlexAlloc: DPM Metadata Organization}
\label{fig:dpmmetadata}
\end{figure}

If the host is under a lot of memory pressure, the host can request that DPM give control
of available regions back to the host. The host does so by invoking a hypercall that passes a page $P$, which DPM will populate with
an array of metadata pertaining to regions that it is handing back. DPM will first unmap this
page $P$ from the host. DPM will then search through the list mentioned above for all available regions and add them
to $P$. Finally, DPM will remap $P$ to the host and return. The host can then
call \textit{free\_pages} on all of the regions listed in $P$.

It is important to ensure that DPM can receive the pages that it needs for its own metadata.
If the host receives an error when trying to hand a region of some size to DPM, the host will know that it needs
to first pass a region of size \textit{PAGE\_SIZE} to DPM in a seperate hypercall, and then retry the
original region pass hypercall, before finally retrying the original hypercall pertaining
to some other EL2 service. 
When DPM receives a region and notices that it is out of metadata
space, it will search for a region of \textit{PAGE\_SIZE} to claim for itself before processing
the larger region. If one is found, then DPM will claim it and process the region hand off
as normal.

\subsubsection{Considerations and Drawbacks}
DPM allows services within EL2 to request arbitrary regions of contiguous memory
at arbitrary points in time. This is valuable when we may or may not need additional pages.
If we know some hypercall will definitely require EL2 to receive more pages,
the error-check-retry pathway may be an unnecessary cost. For example, when we want to
boot a new VM, we know EL2 will definitely need a memory region for the stage-2 page table of the VM.
However, since regions returned to DPM can be re-allocated to other EL2 services as long
as memory is not reclaimed by the host, a new VM could potentially reuse the
region returned by a previously-shutdown VM. DPM is designed to be convenient to use,
but its generalized strategy can lead to extra, avoidable work. We design DPM with
programmability in mind over efficiency, in hopes that it would make full dynamic allocation
attainable for SeKVM services by avoiding excessively-customized handoffs between the host
and corevisor. Although we only partially implement DPM, it would be interesting to
compare the real-world performance metrics between our current stage-2 dynamic allocation
strategy, and a stage-2 dynamic allocation strategy built on top of DPM and a
check-error-retry methodology.

%-------------------------------------------------------------------------------
\section{Conclusions}
%-------------------------------------------------------------------------------

We have designed and implemented FlexAlloc, a system for dynamic memory
partitioning between the corevisor and hostvisor in SeKVM. In particular, we
modify SeKVM so that the hostvisor allocates memory to the corevisor when a VM is
created, and reclaims memory from the corevisor when a VM is destroyed. As a
result, the memory for a VM's stage-2 page table does not need to be statically
allocated to the corevisor upon bootup. This system reduces the wastage of memory
when VM demand is low, and increases the number of VMs that can be run
on a single machine when VM demand is high. We also make a distinct contribution in terms of fragmenting
a VM's stage-2 page table so that it can be composed of multiple noncontiguous
regions of memory, instead of a single contiguous 8M region. Altogether, this
system improves hardware utilization and reduces VM operating costs.

%-------------------------------------------------------------------------------
\section*{Acknowledgments}
%-------------------------------------------------------------------------------

Shih-Wei Li, John S. Koh, and Jason Nieh implemented HypSec, which serves as
the predecessor to SeKVM. Shih-Wei Li, Xupeng Li, Ronghui Gu, Jason Nieh, and
John Zhuang Hui implemented and formally verified SeKVM. Xuheng Li wrote a
helpful set of notes describing the boot and initialization process in SeKVM,
as well as the steps needed to build and run SeKVM~\cite{xuheng}. Jason Nieh and
Xuheng Li provided helpful comments on the project and helped with the design
of FlexAlloc. This paper was written as part of the class COMS E6118 Operating
Systems II at Columbia University.

%-------------------------------------------------------------------------------
\bibliographystyle{plain}
\bibliography{\jobname}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\end{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%